---
title: "Assignment 3"
output: pdf_document
date: "2023-09-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(dplyr)
library(stringr)
```

## #1. Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"

```{r}
url <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv"

majors <- read.csv(file = url, header = T)

str_subset(majors$Major, "DATA|STATISTICS", negate = F)

```

## #2. Write code that transforms the data below:
[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"

**Note: I realized the purpose of this question after I finished questions 3 & 4, so that helped with making the str_extract code and regex**
```{r}
berry_str <- '"bell pepper"  "bilberry"     "blackberry"   "blood orange" "blueberry"    "cantaloupe"   "chili pepper" "cloudberry" "elderberry"   "lime"         "lychee"       "mulberry" "olive"        "salal berry"'

berry_str

berries <- unlist(str_extract_all(berry_str, "\\w+\\s*\\w+"))

berries
```

## #3 Describe, in words, what these expressions will match:
(.)\1\1
"(.)(.)\\2\\1"
(..)\1
"(.).\\1.\\1"
"(.)(.)(.).*\\3\\2\\1"

To do this, I tested a few dataframes in the rstrings package, and then I made my own, consisting of multiple letters and combinations of those letters
```{r}
strings <- c("aaa", "bbb", "ccc", "ddd", "abc", "abcd", "abcde",
             "aaaabbbbccccddddeeee", "alpha", "beta")

long_words <- c("pneumonoultramicroscopicsilicovolcanoconiosis",
                "Pseudopseudohypoparathyroidism", "Floccinaucinihilipilification", "Antidisestablishmentarianism", 
                "Supercalifragilisticexpialidocious",
                "Incomprehensibilities", "Euouae",
                "Unimaginatively", "Honorificabilitudinitatibus",
                "Tsktsk", "Sesquipedalianism")
```
Testing the regex expressions, I determined that:
(.)\1\1 returns strings consisting of three letters repeating in a row

(.)(.)\\2\\1 returns strings containing two characters that are followed by the same characters in reverse

(..)\1 Returns strings containing two letters that are repeated in the same order 

I had to experiment with (.).\\1.\\1 a bit, but it seems to return strings containing a letter followed by a different letter, followed by the same letter as the first letter

(.)(.)(.).*\\3\\2\\1 returns strings containing three letters followed by any other letters, and ending with the first three letters in (I believe) reverse order 

```{r}
str_view(strings, "(.)\\1\\1")
str_view(fruit, "(.)(.)\\2\\1", match = T)
str_view(words, "(..)\\1", match = T)
str_view(long_words,"(.).\\1.\\1", match = T)
str_view(long_words, "(.)(.)(.).*\\3\\2\\1", match = T)
```
## #3 Construct regular expressions to match words that:
Start and end with the same character.
Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)


Note: I used the following website for help with regards to understanding regex syntax: https://www3.ntu.edu.sg/home/ehchua/programming/howto/Regexe.html
```{r}
##Start and end with the same character
str_view(strings, "^(.).*\\1$")

##Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)

str_view(berries, "(\\w+\\w).*\\1")

##Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

str_view(berries, "(\\w).*\\1.*\\1")
```

