---
title: "DATA 607 Week 9 Assignment"
author: "Matthew Roland"
date: "2023-10-25"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("devtools")
#library(devtools)
#devtools::install_github("mkearney/nytimes")
library(tidyverse)
library(dplyr)
library(tidyr)
library(jsonlite)
library(knitr)
```

## Article Search API

For this assignment, I will be using the NYTimes Books API to return a list of all the NYT Best Sellers, and then I will convert the resultant JSON file into an R dataframe.

##Running the API
```{r}
##This code will generte a JSON format list containing historical data pertaining to some of the NYT best sellers from the pat decade (CONFIRM)

nyt <- fromJSON("https://api.nytimes.com/svc/books/v3/lists/best-sellers/history.json?api-key=PCBxRC5IDGa5YL921cADZrhRF0J3B7Bm", flatten = T)

##And now I will convert the JSON list into a dataframe
nyt <- data.frame(nyt)

kable(nyt)

```

 As we can see, this generates a rather unorganized (and incomplete) dataframe consisting of empty and nested columns (and even empty, nested columns)
Of course, it is necessary to clean and tidy all of these datapoints

```{r}
##First, I want to remove the "results." prefix

colnames(nyt) <- sub("^results.", "", colnames(nyt))

nyt_unnest <- nyt %>% unnest_wider(c("ranks_history", "reviews", "isbns"))

#dropping the primary isbn variables, since they contain the same information contained within the nested isbn variable (and the nested variable seems to contain more complete observations)

nyt_unnest <- subset(nyt_unnest, select = -c(primary_isbn10, primary_isbn13))


nyt_unnest <- nyt_unnest %>% unnest_longer(isbn10)
nyt_unnest <- nyt_unnest %>% unnest_longer(isbn13)
nyt_unnest <- nyt_unnest %>% unnest_longer(rank)
nyt_unnest <- nyt_unnest %>% unnest_longer(list_name)
nyt_unnest <- nyt_unnest %>% unnest_longer(display_name)
nyt_unnest <- nyt_unnest %>% unnest_longer(published_date)
nyt_unnest <- nyt_unnest %>% unnest_longer(bestsellers_date)
nyt_unnest <- nyt_unnest %>% unnest_longer(weeks_on_list)
nyt_unnest <- nyt_unnest %>% unnest_longer(rank_last_week)
nyt_unnest <- nyt_unnest %>% unnest_longer(asterisk)
nyt_unnest <- nyt_unnest %>% unnest_longer(dagger)

glimpse(nyt_unnest)

nyt_unnest[nyt_unnest == ""] <- NA

```

And with that, we are left with a very, very large (albeit analyzable) unnested dataset
